---
title: "Regression Models Final Project"
author: "Fernando Gonzalez Prada"
date: "21 de agosto de 2015"
output: html_document
---

***"Is an automatic or manual transmission better for MPG?"***

***"Quantify the MPG difference between automatic and manual transmissions?"***

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

A data frame with 32 observations on 11 variables.  


* **mpg:** Miles/(US) gallon
* **cyl:** Number of cylinders
* **disp:** Displacement (cu.in.)
* **hp:** Gross horsepower
* **drat:** Rear axle ratio
* **wt:** Weight (lb/1000)
* **qsec:** 1/4 mile time
* **vs:** V/S
* **am:** Transmission (0 = automatic, 1 = manual)
* **gear:** Number of forward gears
* **carb:** Number of carburetors 


Open the data
```{r Loading the data}
suppressMessages(library(datasets))
suppressMessages(library(pastecs))
suppressMessages(library(ggplot2))

df <- mtcars
df$am <- factor(df$am, labels = c("Automatic", "Manual"))
summary(df)
```

**Simple Linear Regression Model**
```{r simple model}
simple <- lm(mpg ~ am, data = df)
summary(simple)
```

**Multiple Linear Regression Model and Stepwise**
```{r step1}
multi1 <- step(
                lm(mpg ~ ., data = df)
                    , trace = 0, steps = 10000, direction = "both");
summary(multi1)
```


**Scale and Center the numeric variables**

```{r center and scale}

df$mpg = scale(df$mpg, center = TRUE, scale = TRUE)
df$disp = scale(df$disp, center = TRUE, scale = TRUE)
df$hp = scale(df$hp, center = TRUE, scale = TRUE)
df$drat = scale(df$drat, center = TRUE, scale = TRUE)
df$wt = scale(df$wt, center = TRUE, scale = TRUE)
df$qsec = scale(df$qsec, center = TRUE, scale = TRUE)
df$vs = scale(df$vs, center = TRUE, scale = TRUE)
df$gear = scale(df$gear, center = TRUE, scale = TRUE)
df$carb = scale(df$carb, center = TRUE, scale = TRUE)
# summary(df)
```

**Multiple Linear Regression Model and Stepwise, with centered and scaled variables**
```{r multi2}
multi2 <- step(
                lm(mpg ~ ., data = df)
                    , trace = 0, steps = 10000, direction = "both");
summary(multi2)
```

**Anova Analisys
```{r Anova}
anova(simple, multi1, multi2)
```


** Apendix

**Checking for normality:**

*Perform Shapiro Test*
 
```{r Shapiro}
round(
    stat.desc(
        df[, c("mpg", "disp", "hp", "drat", "wt", "qsec", "vs", "gear", "carb")],
        basic = FALSE, norm = TRUE, desc=FALSE)
    ,digits = 3)
```
Observing the values of the "normtest.W" row, most of the variables have values close to 1, which is an indicator the the data is nearly normally distributed.


*The quantile plots also show that the columns have normal distribution:*

```{r Qplots, echo=FALSE}

qqplot.mpg <- qplot(sample = df$mpg, stat="qq") + labs(x = "mpg")
qqplot.mpg
qqplot.disp <- qplot(sample = df$disp, stat="qq") + labs(x = "disp")
qqplot.disp
qqplot.hp <- qplot(sample = df$hp, stat="qq") + labs(x = "hp")
qqplot.hp
qqplot.drat <- qplot(sample = df$drat, stat="qq") + labs(x = "drat")
qqplot.drat
qqplot.wt <- qplot(sample = df$wt, stat="qq") + labs(x = "wt")
qqplot.wt
qqplot.qsec <- qplot(sample = df$qsec, stat="qq") + labs(x = "qsec")
qqplot.qsec
qqplot.gear <- qplot(sample = df$gear, stat="qq") + labs(x = "gear")
qqplot.gear
qqplot.carb <- qplot(sample = df$carb, stat="qq") + labs(x = "carb")
qqplot.carb
```


